Distributional Trees for Circular Data

### 2019-02-04 pkg/disttree/inst/IWSM2019/abstract_IWSM2019.tex
further revision
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@2431 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: bae0c8c64bf67fb2fa011fce6b8a27ad432ab088

\section{Discussion and Outlook}
This study shows that applying distributional trees on circular data provides easily 
interpretable models clearly illustrating the impact of given covariates and their possible 
interactions on the distribution of the dependent variable. 
Moreover, covariates are selected automatically such that no prespecification of the model is 
required which can be a challenging task, particularly for a meteorlogical quantity such as wind 
direction with high variation and many possible influencing factors.
%Another valuable feature is the capability of capturing also non-additive and abrupt changes, 
Additionally, abrupt changes are captured, e.g., when the impact of a covariate such as 
\textit{air pressure difference} determines either up- or down-valley wind directions.

A possible extension of the method as presented is to build an ensemble of 
distributional trees to regularize and stabilize the model 
(see, e.g., distributional forests, Schlosser et.~al, 2018). 
%In that way, it should not have a 
%significant impact on the results whether circular covariates are included as linear regressors
%yielding a larger number of splits or as circular regressors with ``intervall splits''.

Note that the circular variable $dd$ (lagged observations of wind) is employed as a linear
regressor. Therefore, several splits are performed in $dd$ in order to properly cluster the training 
data in corresponding wind directions (see Figure~\ref{fig:tree}). 
In this case searching for two split points defining a split interval on the circular scale instead 
of a single split point on the linear scale could reduce the complexity of the model and possibly 
improve the results.
%Alternatively, in the case of a circular covariate 
%the tree building algorithm could search for two split points defining a split interval instead of 
%a single split point. This would reduce the complexity of the model and possibly improve the results.
%Based on the results it can be assumed that including covariates such as lagged observations of 
%wind direction as linear regressors provides valuable information. 
%However, in order to properly account not only for the circular properties of the response but 
%also of regressors, additional adaptions of distributional trees are required. In particular, 
%if the selected split variable is a circular variable, e.g. wind direction, this demands for finding
%two split point defining a ``split interval'' instead of a single split point.
%This is also supported by the results illustrated in Figure~\ref{fig:tree} as mutliple splits in 
%the linearly employed regressor \textit{dd} (lagged wind direction) are required in order to 
%cluster the training data in corresponding wind directions.  
\textit{Computational Details:} The employed \textsf{R} packages are available on \textsf{R}-Forge at
https://R-Forge.R-project.org/projects/partykit/.

>>>>

\section{Discussion and Outlook}
This study shows that applying distributional trees on circular data provides easily 
interpretable models clearly illustrating the impact of given covariates and their possible 
interactions on the distribution of the dependent variable. 
Moreover, covariates are selected automatically such that no prespecification of the model is 
required which can be a challenging task, particularly for a meteorlogical quantity such as wind 
direction with high variation and many possible influencing factors.
Additionally, abrupt changes are captured, e.g., when the impact of a covariate such as 
\textit{air pressure difference} determines either up- or down-valley wind directions.

A possible extension of the method as presented is to build an ensemble of 
distributional trees to regularize and stabilize the model 
(see, e.g., distributional forests, Schlosser et.~al, 2018). 
% In that way the impact of the chosen scale on which circular regressors are considered should be reduced.

Note that circular variables are employed as linear regressors which leads to multiple splits
in these variables in order to properly cluster the training data.
%Therefore, considering them on a circular scale could reduce the complexity of the model and
%possibly improve the results.
%Note that the circular variable $dd$ (lagged observations of wind) is employed as a linear
%regressor. Therefore, several splits are performed in $dd$ in order to properly cluster the training 
%data in corresponding wind directions (see Figure~\ref{fig:tree}). 
In this case searching for two split points defining a split interval on the circular scale instead 
of a single split point on the linear scale could reduce the complexity of the model and possibly 
improve the results.
%Alternatively, in the case of a circular covariate 
%the tree building algorithm could search for two split points defining a split interval instead of 
%a single split point. This would reduce the complexity of the model and possibly improve the results.
%Based on the results it can be assumed that including covariates such as lagged observations of 
%wind direction as linear regressors provides valuable information. 
%However, in order to properly account not only for the circular properties of the response but 
%also of regressors, additional adaptions of distributional trees are required. In particular, 
%if the selected split variable is a circular variable, e.g. wind direction, this demands for finding
%two split point defining a ``split interval'' instead of a single split point.
%This is also supported by the results illustrated in Figure~\ref{fig:tree} as mutliple splits in 
%the linearly employed regressor \textit{dd} (lagged wind direction) are required in order to 
%cluster the training data in corresponding wind directions.

\medskip

\textit{Computational Details:} The employed \textsf{R} packages are available on \textsf{R}-Forge at
https://R-Forge.R-project.org/projects/partykit/.
